CF Registry Analysis

### 2022-12-27 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd

```{r}
# All data
encounter <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
                      na.strings = c("", "N/A")
)
annual <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
                   na.strings = c("", "N/A")
)
encounter <- encounter %>%
  arrange(eDWID, reviewyear, encounternum)
annual = annual %>% arrange(eDWID,ReviewYear)
# Add age
ages <- annual %>% 
  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd)) %>%
  rename(reviewyear = ReviewYear) %>%
  group_by(eDWID,reviewyear) %>%
  summarise(`Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), 
                                                  decreasing = TRUE)[1])),
            Age = Age_YrEnd[1],.groups = "drop")
encounter = left_join(encounter,ages,by = c("eDWID", "reviewyear"))
encounter$`Age Group 2018` = cut(encounter$`Age in 2018`,
                            c(-Inf,10,19,31,Inf),
                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
                            right=F)
encounter$`Age Group` = cut(encounter$Age,
                            c(-Inf,10,19,31,Inf),
                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
                            right=F)
# For those with CFRD by OGTT, find first visit with diagnosis
first_cfrd_ogtt <- encounter %>% 
  filter(twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_ogtt = reviewyear)
# For those with CFRD by registry criteria, find first visit with diagnosis
first_cfrd_cffpr <- encounter %>% 
  filter(cfrd_status==3) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_cffpr = reviewyear)
# Both
first_cfrd_both <- encounter %>% 
  filter(cfrd_status==3|twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_both = reviewyear)
# Add to encounter level data
encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age in 2018"))
encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age in 2018"))
encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age in 2018"))
# Fill down CFRD Yes/No by year
encounter <- encounter %>%
  group_by(eDWID) %>%
  mutate(
    cfrd_ogtt = case_when(
      reviewyear < first_cfrd_ogtt ~ "No",
      reviewyear >= first_cfrd_ogtt ~ "Yes"
    ),
    cfrd_cffpr = case_when(
      reviewyear < first_cfrd_cffpr ~ "No",
      reviewyear >= first_cfrd_cffpr ~ "Yes"
    ),
    cfrd_both = case_when(
      reviewyear < first_cfrd_both ~ "No",
      reviewyear >= first_cfrd_both ~ "Yes"
    )
  ) %>%
  ungroup()
encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
encounter$cfrd_both[is.na(encounter$cfrd_both)] <- "No"
# Ordered factors
encounter$cfrd_ogtt = factor(encounter$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
encounter$cfrd_cffpr = factor(encounter$cfrd_cffpr,levels = c("No","Yes"),ordered = T)
encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
# Remove those missing age
encounter = encounter %>% filter(!is.na(Age))
```

>>>>

```{r}
# All data
encounter <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
                      na.strings = c("", "N/A")
)
annual <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
                   na.strings = c("", "N/A")
)
encounter <- encounter %>%
  arrange(eDWID, reviewyear, encounternum)
annual = annual %>% arrange(eDWID,ReviewYear)
# Add age
ages <- annual %>% 
  rename(reviewyear = ReviewYear) %>%
  group_by(eDWID,reviewyear) %>%
  summarise(Age = as.numeric(Age_YrEnd[1]),.groups = "drop")
encounter = left_join(encounter,ages,by = c("eDWID", "reviewyear"))
encounter$`Age Group` = cut(encounter$Age,
                            c(-Inf,10,19,31,Inf),
                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
                            right=F)
# For those with CFRD by OGTT, find first visit with diagnosis
first_cfrd_ogtt <- encounter %>% 
  filter(twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age Group`) %>%
  rename(first_cfrd_ogtt = reviewyear)
# For those with CFRD by registry criteria, find first visit with diagnosis
first_cfrd_cffpr <- encounter %>% 
  filter(cfrd_status==3) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age Group`) %>%
  rename(first_cfrd_cffpr = reviewyear)
# Both
first_cfrd_both <- encounter %>% 
  filter(cfrd_status==3|twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age Group`) %>%
  rename(first_cfrd_both = reviewyear)
# Add to encounter level data
encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age Group"))
encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age Group"))
encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age Group"))
# Fill down CFRD Yes/No by year
encounter <- encounter %>%
  group_by(eDWID) %>%
  mutate(
    cfrd_ogtt = case_when(
      reviewyear < first_cfrd_ogtt ~ "No",
      reviewyear >= first_cfrd_ogtt ~ "Yes"
    ),
    cfrd_cffpr = case_when(
      reviewyear < first_cfrd_cffpr ~ "No",
      reviewyear >= first_cfrd_cffpr ~ "Yes"
    ),
    cfrd_both = case_when(
      reviewyear < first_cfrd_both ~ "No",
      reviewyear >= first_cfrd_both ~ "Yes"
    )
  ) %>%
  ungroup()
encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
encounter$cfrd_both[is.na(encounter$cfrd_both)] <- "No"
# Ordered factors
encounter$cfrd_ogtt = factor(encounter$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
encounter$cfrd_cffpr = factor(encounter$cfrd_cffpr,levels = c("No","Yes"),ordered = T)
encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
# Remove those missing age
encounter = encounter %>% filter(!is.na(Age))
```
